Reconstructing Propositional Proofs in Type Theory

### 2017-10-23 paper/sections/metis-language-and-proofs.tex
[ paper ] fixed minor typos.

@@

@@

AGREGAR EL TEXTO

% \TSTP derivations replied by \Metis use the following rules:
% \canonicalize, \clausify, \conjunct, \simplify, \negate, and \strip.

@@

A proof delivered by \Metis encodes a natural
deduction proof (see a complete proof in Fig.~\ref{fig:metis-proof-tstp}).
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (\eg,
\emph{falsium} in the root of the \TSTP derivation
in Fig.~\ref{fig:metis-proof-tstp}).

These proofs are directed acyclic graphs, trees of refutations
(see the proof tree of the \TSTP derivation showed above in
Fig.~\ref{fig:metis-proof-tstp}).
Each node stands for an application of an inference rule and the leaves
in the tree represent formulas in the given problem. Each node is
labeled with a name of the inference rule (e.g., \canonicalize).
Each edge links a premise with one conclusion.
The proof graphs have at their root the conclusion
$⊥$, since \Metis delivers proofs by refutation.

>>>>

\begin{remark}
We address proof-reconstruction in this paper for refutation proofs.
We annotate \abbre{CPL} problems using \verb!fof! formulas that
\Metis converts to \abbre{CNF} clauses to look for a refutation.
\end{remark}

@@

A proof delivered by \Metis encodes a natural
deduction proof.
With these inference rules as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (\eg,
\emph{falsium} in the root of the \TSTP derivation).

These proofs are directed acyclic graphs, trees of refutations
Each node stands for an application of an inference rule and the leaves
in the tree represent formulas in the given problem. Each node is
labeled with a name of the inference rule (\eg, \canonicalize).
Each edge links a premise with one conclusion.
The proof graphs have at their root the conclusion
$⊥$, since \Metis delivers proofs by refutation.

@@

In the following table we present the list of inference rules used
by \Metis in its \TSTP derivations for propositional logic.

@@

### 2017-10-23 paper/sections/proof-reconstruction/strip.tex
[ paper ] fixed rm funnctions and example added.
[ paper ] fixed minor typos.

We now define the $\fstripp$ function in~(\ref{eq:strip-unfixed}) that
yields the conjunction of subgoals that implies the goal of the problem in
the \Metis \TSTP derivations.

@@

%  1137  git add .
%  1138  git commit -am "[ paper ] strip proof."
%  1139  git pull origin paper
%  1140  git push origin paper
\begin{equation}
\label{eq:strip-fixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{llll}
\fstrip &(φ₁ ∧ φ₂)     &(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     &(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      &(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         &(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         &(\suc~n) &= ⊥\\
\fstrip &φ             &\text{n} &= φ
\end{array}
\end{aligned}
\end{equation}

@@

\begin{mainth}
\label{thm:strip}
If $s_1$ and $s_2$ are the subgoals of the goal $φ$, that is,
$\fstrip~φ~n~≡~s_1\wedge~s_2$, if $Γ ⊢ s_1$ and $Γ ⊢ s_2$ then $Γ ⊢ φ$.
\end{mainth}

>>>>

We now define the $\fstripp$ function in~(\ref{eq:strip-unfixed}) that
yields the conjunction of subgoals that implies the goal of the problem in
the \Metis \TSTP derivations. Its definition is based on the source code
of \Metis for the strip goal functions.

@@

%  1137  git add .
%  1138  git commit -am "[ paper ] strip proof."
%  1139  git pull origin paper
%  1140  git push origin paper

\begin{equation}
\label{eq:strip-fixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{llll}
\fstrip &(φ₁ ∧ φ₂)     &(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     &(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      &(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         &(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         &(\suc~n) &= ⊥\\
\fstrip &φ             &\text{n} &= φ
\end{array}
\end{aligned}
\end{equation}

@@

\begin{mainth}
\label{thm:strip}
Let $\text{n} : \Nat$ be the complexity measure of the strip function defined
in~(\ref{eq:strip-fixed}).
If $s_1$ and $s_2$ are the subgoals of the goal $φ$, that is,
$$\fstrip~φ~n~≡~s_1∧~s_2,$$
if $Γ ⊢ s_1$ and $Γ ⊢ s_2$ then $Γ ⊢ φ$.
\end{mainth}
